---
title: "04_fit_models.Rmd"
author: "Dexter H. Locke, PhD"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assumes "02_Woosley_combine_data.Rmd" was run



## 0 set up: load libraries, custom functions, set defaults
```{r}
# load libraries
# packages we'll be using
packs <- c('tidyverse'        # a must have!
           , 'tidylog'        # makes things very verbose for 2x checking 
           , 'magrittr'       # all of the pipes
           , 'janitor'        # cleans things up
           # , 'sf'             # simple features
           # , 'mapview'        # quick webmaps for zoom/pan viz
           #'tidycensus',     # access to Census data in a tidy way
           #'party',          # random forests
           , 'modEvA'         # contains D-squared function
           , 'randomForest'  #'rfUtilities', 'verification'
           , 'tictoc'         # times things
           , 'beepr'          # makes noises
           # , 'broom'          # tidy up regression models
           # , 'performance'    # nice regression diagnostics
           # , 'psych'          # describe is very useful for descriptive statistics
           # , 'sjPlot'         # useful plotting and regression support
           # , 'rpart'          # for random forests
           # # , pROC           # for AUC calculation
           , 'pdp'            # partial dep plots
           , 'vip'            # Variable Importance Plots (and extracting vip vals)
           # , 'caret'          #random forest work
           , 'party'          # for random forests using cforest
           )        

# check for all of the libraries
if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))  
}

# lapply(packs, library, character.only = TRUE)
vapply(packs, library, character.only = TRUE, logical(1),
       logical.return = TRUE, quietly = TRUE)


# custom function for "Not In"
`%nin%` <- Negate(`%in%`)

# for reproducibility, we should have the same random draws. setting the seed ensures that is the case.
set.seed(19870630)
```



## 1 read in data
### A main `build` file
```{r}

# one 'main' data frame
build <- read_csv(paste0(getwd(), '/output_data/building_2022-02-01.csv')) %>%
  tidylog::select(
    # damage_binary
      damage_cat         # binary outcome encoded as categorical (Destroyed / Survived)
    , damage_severity    # five levels
    , build_area   
    # build_DECKPORCHE    , #drop this on 12/20/21
    , build_DECKPORCHO
    , build_EAVES
    , build_EXTERIORSI
    , build_FENCEATTAC
    , build_PATIOCOVER
    , build_PROPANETAN
    , build_ROOFCONSTR
    , build_VEGCLEARAN 
    , build_VENTSCREEN 
    , build_WINDOWPANE 
    , parcel_year_built 
    , build_near_build 
    , build_has_tree_overhang 
    , build_min_dist_shrub 
    , build_min_dist_tree 
    , build_overhang_ht 
    , build_perc_overhang 
    , parcel_Build_P 
    , parcel_Can_P 
    , parcel_Grass_P 
    , parcel_Imperv_P    #inverse of perv surface
    , parcel_Paved_P    
    #  parcel_Perv_P    ,#inverse of imp, just need one
    , parcel_Road_P
    , parcel_Shrub_P  
    , parcel_Soil_P  
    , parcel_Water_P  
    , build_near_dest  
    , build_p_building_10  
    , build_p_building_100  
    , build_p_building_200  
    , build_p_building_300  
    , build_p_grass_10  
    , build_p_grass_100  
    , build_p_grass_200  
    , build_p_grass_300  
    , build_p_otherpaved_10  
    , build_p_otherpaved_100  
    , build_p_otherpaved_200  
    , build_p_otherpaved_300  
    , build_p_road_10  
    , build_p_road_100  
    , build_p_road_200  
    , build_p_road_300  
    , build_p_shrub_10  
    , build_p_shrub_100  
    , build_p_shrub_200  
    , build_p_shrub_300  
    , build_p_soil_10  
    , build_p_soil_100  
    , build_p_soil_200  
    , build_p_soil_300  
    , build_p_tree_10  
    , build_p_tree_100  
    , build_p_tree_200  
    , build_p_tree_300  
    , build_p_water_10  
    , build_p_water_100  
    , build_p_water_200  
    , build_p_water_300  
    , parcel_area  
    , build_Mean_builddens  
    , build_Mean_distall_road  
    , build_Mean_distroad  
    , build_Mean_aspect_100m_DEM  
    , build_Mean_aspect_30m_DEM  
    , build_Mean_elev_100m  
    , build_Mean_elev_30m  
    , build_Mean_slope_100m_DEM  
    , build_Mean_slope_30m_DEM
    , build_DINS
    ) %>%
  mutate_if(is.character, as.factor) %>% # turns characters into factors
  droplevels() %>% 
  data.frame() # some of the random forest functions do not 'like' tibbles
               # since the non-dins and dins data will be pulled from here
               # we change here so that the subsets are already formatted as data.frames

# Miranda: do these NA's surprise you? OK to drop non-DINS-associated NA's?
# where are the NA's
build %>% 
  map(~sum(is.na(.))) %>% # map is like for loop but vectorized, meaning its faster uses dif. syntax help(purrr::map)
  bind_rows() %>%         # the output of purrr::map is a list, this makes a tibble
  t() # %>% View() # check for NA's ("t" is for Transpose, just makes it easier to read)

# Miranda, are these NA's expected?
```



### B non-DINS data (larger dataset, fewer predictors)
```{r}

# a subset for non-DINS-associated data
build_no_dins <- build %>% 
  tidylog::select(# dropping columns containing specific DINS data
                  # very detailed but sparse (lots of NAs)
    !c(  build_DECKPORCHO
       , build_EAVES
       , build_EXTERIORSI
       , build_FENCEATTAC
       , build_PATIOCOVER
       , build_PROPANETAN
       , build_ROOFCONSTR
       , build_VEGCLEARAN
       , build_VENTSCREEN
       , build_WINDOWPANE
       , build_DINS
       , damage_severity # dropping the DV for the smaller model
       )
    )

# Miranda, are these NA's expected? Should we drop them or find a random forest model that can make use of missing
# data?
build_no_dins %>% 
  map(~sum(is.na(.))) %>% 
  bind_rows() %>% 
  t() # %>% View() # check for NA's

# how is the outcome distributed?
build         %>% tabyl(damage_cat) # good double check
build_no_dins %>% tabyl(damage_cat)

```



### C DINS data (smaller dataset, more predictors)
```{r}
# get data for DINS analyses
build %>% tabyl(build_DINS)

# subset
build_dins <- build %>% 
  filter(build_DINS == 1) %>% 
  select(-build_DINS, -damage_cat)
  # dropping the DV for the smaller model

# Miranda, even after dropping the non-DINS data are there are a lot of NA's in the DINS columns
# do we need all of them?
# in other words, not even the DINS data are complete for DINS data
build_dins %>% 
  map(~sum(is.na(.))) %>% 
  bind_rows() %>% 
  t() # %>% View() # check for NA's


# when we keep only complete cases (so remove all NA's) the file is really short (n = 172)
# is this a surprise?
dim(build_dins)
build_dins %>% drop_na() %>% dim() # not going to be practical for randomForest::randomForest

build_dins %>% tabyl(damage_severity)

```



## 2 fit random forest models
### A non-DINS analyses (larger dataset, fewer predictors)
```{r}
# fit random forest models
# randomForest::randomForest is faster but requires full data
# party::cforest is slower but can handle missing data

n_tree <- 10 # used for setting number of trees
n_tree <- 500 

# random forest 
tic(); rf_no_dins <- randomForest::randomForest(
  damage_cat ~ .
  # need full dataset, can't keep NA's
  , data = build_no_dins %>% drop_na()
  , type = 'classification' # not regression
  , mtry = 5
  , ntree = n_tree # 10 ~.5 seconds,100 ~5s, 200 ~7.5s, 3 ~12s, ... 5 ~23s 
                   # tested on a few smaller forests
  ); toc()


# # since this takes a long time to run, I've commented the code off and just read-in
# # the completed model file in its "*.Rdata" format with the load() function.
# # random forest via party::cforest which accepts missing data and is slower 
# tic(); cf_no_dins <- party::cforest(
#   damage_cat ~ .
#   , data = build_no_dins # notice no drop_na here
#   , controls =
#     cforest_control(
#         mtry = 5
#       , ntree = n_tree # 10 ~2 seconds, 100 ~25s, 200 ~50s, 300 ~75s ... 500 ~130 seconds
#       , mincriterion = 0
#       )
#   ); toc()
# 
# # save out since it is slow
# tic(); save(cf_no_dins
#             , file = paste0('../Massive_forests/cf_no_dins_', Sys.Date(), '.Rdata')); toc()

# read back in
tic(); load(file = '../Massive_forests/cf_no_dins_2022-02-22.Rdata'); toc() # ~9s

```


### B DINS analyses (smaller dataset, more predictors)
```{r}

# predicting severity

# random forest via party::cforest which accepts missing data and is slower
tic(); cf_dins <- party::cforest(
  damage_severity ~ . # severity, not binary
  , data = build_dins # notice build_dins here, no "no" in the object name.
  , controls =
    cforest_control(
        mtry = 5
      , ntree = n_tree # 10 ~0.52 seconds, 100 ~1.5s, 200 ~2.7s, 300 ~3.5s ... 500 ~7s
      , mincriterion = 0
      )
  ); toc()

# not sure if saving is needed..
tic(); save(cf_dins
            , file = paste0('../Massive_forests/cf_dins_', Sys.Date(), '.Rdata')); toc() # 3.5 seconds

# read back in
tic(); load(file = '../Massive_forests/cf_dins_2022-02-22.Rdata'); toc() # 1.5 seconds


```


## 3 overall model accuracy
under construction :-/ DHL todo
"100_old_hold.Rmd" has good code to pull from

also the ROC curves from "### v how did the model perform?" chunk in "Philly_Land_Cover_Change.Rmd"


## 4 variable importance
```{r}
# # modified from first chunk here:
# # https://bgreenwell.github.io/pdp/articles/pdp.html#other-vignettes
# # Variable importance plot (compare to randomForest::varImpPlot(boston_rf))
# vip(boston_rf, bar = FALSE, horizontal = FALSE, size = 1.5)  # Figure 1

# variable importance plot with random forest; fast
vip_rf_no_dins <- vip(rf_no_dins
    # , bar = FALSE       # doesn't seem to work "help(vip)" showed that it changed to "geom"
    # , geom = 'point'    # try "col" instead
    , geom = 'col'      # try "point" instead
    , horizontal = TRUE # vignette example had this as FALSE (label easier rotated)
    # , size = 1.5        # plausibly pertains to point size
    # see "num_features"
    # "Integer specifying the number of variable importance scores to plot. Default is 10."
    ) + 
  theme_bw(16)          # added some flare (ok just bigger axis text and black/white theme)

# check the plot
vip_rf_no_dins



# if we want to use color/shapes or facets to indicate the variable TYPE
# like we did with the cforest object before, it looks like vip::vi can be used
# to calculate the variable importance scores, then we can re-purpose the old
# code to add the colors symbols, etc.. . 'help(vip)' has an example starting with 
# "# Better yet, store the variable importance scores and then plot
# # vi_scores <- vi(model, method = "firm")". That's how I learned to extract importance
# from randomForest::randomForest models

# # since the second one takes 10 minutes, this code is commented out. Instead of re-running, just read a *.csv backin
# # not sure about this scaling. When scale is FALSE the values are on totally different scales
# # `rank = TRUE`
# tic(); vi_rf_no_dins <- vi(rf_no_dins, scale = TRUE) %>% mutate(Model = 'rf_no_dins', n_tree = n_tree); toc()
# tic(); vi_cf_no_dins <- vi(cf_no_dins, scale = TRUE) %>% mutate(Model = 'cf_no_dins', n_tree = n_tree); toc() # 10 mins
# tic(); vi_cf_dins    <- vi(cf_dins,    scale = TRUE) %>% mutate(Model =    'cf_dins', n_tree = n_tree); toc() # 1 min
# 
# 
# bind_rows(vi_rf_no_dins, vi_cf_no_dins, vi_cf_dins) %>% # combines variable importance from each model
#   write_csv(paste0('../Massive_forests/variable_importance_', Sys.Date(), '.csv'))

# TODO join in aliases and type, repurpose directly below.
(
  variable_importance <- read_csv('../Massive_forests/variable_importance_2022-02-22.csv') %>% 
    group_by(Model) %>% 
    mutate(rank = row_number(Model)) %>% 
    ungroup()
  )

# # get tab_lab
# left_join(
#       tab_labs %>% 
#         tidylog::select(`Variable name`, Alias, Type, Scale)) 

```


## 5 partial dependence plots
### A base-style graphs (example only)
```{r eval=FALSE, include=FALSE}
# single base-graphics partial dependency plot, modeled after Figure 2 here:
# https://bgreenwell.github.io/pdp/articles/pdp.html#other-vignettes
# partialPlot(boston_rf, pred.data = boston, x.var = "lstat")  # Figure 2
randomForest::partialPlot(rf_no_dins
                          , pred.data = build_no_dins
                          , x.var = "build_near_dest"
                          ) # ugly!
```


### B ggplot graphs
#### i test on a single predictor
```{r}
# # building off of the third chunk from
# # https://bgreenwell.github.io/pdp/articles/pdp.html#other-vignettes
# # AND third chunk from this vignette
# # https://bgreenwell.github.io/pdp/articles/pdp-computation.html
# # Switch to ggplot2
# p2 <- partial(boston_rf, pred.var = "lstat", plot = TRUE,
#               plot.engine = "ggplot2")

# THIS IS A SINGLE VARIBLE TEST
pdp::partial(rf_no_dins                      # random forest model
             , pred.var = "build_near_dest"  # predictor variable
             , plot = TRUE                   # if FALSE returns prediction data
             , plot.engine = "ggplot2"       # plot with ggplot2 or not?
             , rug = TRUE                    # adds the univariate y-axis summary (aka "rug")
             , progress = "text"             # prints progress, sets users expectations
             , quantiles = TRUE              # see help(pdp)
             , probs = 0:20/20               # probabilities, "0:20/20" gives 5% increments
             , ice = TRUE                    # Individual Conditional Expectations
             , center = TRUE                 # for graphical layout, pertains to ice
             , alpha = .1                    # line transparency
             ) + 
  theme_bw(16) +                             # ggplot adjustment for black/white & large text size 
  # scale_y_continuous(expand = c(0, 0)) +
  NULL -> test_pdp                           # makes the plot

test_pdp # accesses plot

variable <- "build_near_dest"
ggsave(file = paste0(getwd(), '/figures/cf_dins/', variable, '_', Sys.Date(), '.png')
       , width = 6.5*2
       , height = 6.5*1.5
       , dpi = 100)
```

#### ii non-DINS RF
```{r}
# TODO
```

#### iii non-DINS CF
```{r}
# TODO Miranda, we can look at all plots, but maybe we just want to look at the top 10 or 20 top predictors

## CF non-DINS
# loop through predictors to get PDP
# first get the iterator, here the variable names

# using ALL predictors
(pdp_cf_no_dins_vars <- variable_importance %>% filter(Model == 'cf_no_dins') %>%                 pull(Variable))

# using top n predictors
top_n <- 10          # set n (can revise to 20 or any other value greater than zero and less than 71)
(pdp_cf_no_dins_vars <- variable_importance %>% filter(Model == 'cf_no_dins') %>% slice(1:top_n) %>% pull(Variable)) 


resolution <- 100                       # dots per inch (dpi) for figure output (lower values keep files size down,
                                        # can increase for publication)
tic()                                   # clock in for whole loop
for(variable in pdp_cf_no_dins_vars){   # start loop, iterating over pdp_cf_no_dins_vars
  print(variable)                       # just to track the progress
  tic()                                 # clock in within the loop
  pdp::partial(cf_dins                  # the model
             , pred.var = variable  
             , plot = TRUE              
             , plot.engine = "ggplot2"  
             , rug = TRUE               
             , progress = "text"        
             , quantiles = TRUE         
             , probs = 0:20/20          
             , ice = TRUE               
             , center = TRUE            
             , alpha = .1               
             ) + 
    theme_bw(16) +                       # ggplot adjustment for black/white & large text size 
    # scale_y_continuous(expand = c(0, 0)) +
    labs(title = variable) +
    NULL
  
  # save out
  ggsave(file = paste0(getwd(), '/figures/cf_no_dins/', variable, '_', Sys.Date(), '.png')
         , width = 6.5*2
         , height = 6.5*1.5
         , dpi = resolution)
  toc()    # clock out within loop
  }; toc() # clock out whole loop



```

#### iv DINS CF
```{r}
# TODO Miranda, we can look at all plots, but maybe we just want to look at the top 10 or 20 top predictors


## CF DINS
# loop through predictors to get PDP
# first get the iterator, here the variable names

# using ALL predictors
(pdp_cf_dins_vars <- variable_importance %>% filter(Model == 'cf_dins') %>%                 pull(Variable))

# using top n predictors
top_n <- 10          # set n (can revise to 20 or any other value greater than zero and less than 71)
(pdp_cf_dins_vars <- variable_importance %>% filter(Model == 'cf_dins') %>% slice(1:top_n) %>% pull(Variable)) 


resolution <- 100                       # dots per inch (dpi) for figure output (lower values keep files size down,
                                        # can increase for publication)
tic()                                   # clock in for whole loop
for(variable in pdp_cf_dins_vars){      # start loop
  print(variable)                       # just to track the progress
  tic()                                 # clock in within the loop
  pdp::partial(cf_dins                  # the model
             , pred.var = variable  
             , plot = TRUE              
             , plot.engine = "ggplot2"  
             , rug = TRUE               
             , progress = "text"        
             , quantiles = TRUE         
             , probs = 0:20/20          
             , ice = TRUE               
             , center = TRUE            
             , alpha = .1               
             ) + 
    theme_bw(16) +                       # ggplot adjustment for black/white & large text size 
    # scale_y_continuous(expand = c(0, 0)) +
    labs(title = variable) +
    NULL
  
  # save out
  ggsave(file = paste0(getwd(), '/figures/cf_dins/', variable, '_', Sys.Date(), '.png')
         , width = 6.5*2
         , height = 6.5*1.5
         , dpi = resolution)
  toc()    # clock out within loop
  }; toc() # clock out whole loop



```




```{r, citations}
lapply(packages, citation)
```


Last knit on `r format(Sys.time())`


```{r}
system.time(save.image(file = paste0('saved_sessions/wui_r_models_', gsub('[[:punct:]]', '-', Sys.time()), '.RData')))
```

